Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change

### 2018-09-27 LatexNotes/Notes.tex
NM

\subsubsection{Non-dimensionalization}

The KdV equation in dimensional form is
\begin{equation}
2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}
Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$. I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
\begin{equation}
2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}
Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.

@@

Let's establish some basic scales and notation:
% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table.}
\label{scale_table}
\begin{tabular}{l l l l}
\hline Description & Notation & Upstream  value \,\, & Downstream value \\
\hline
Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
Depth					& $h_0^{\pm}$			& 12.5 cm		& 3 cm \\
Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

>>>>

\subsubsection{The constant-depth KdV equation in dimensional form}

The constant-depth KdV equation in dimensional form is
\begin{equation}
2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}
Here, $\eta$ has units of length, $t$ time, $h_0$ is the depth, $c = \sqrt{g h_0}$ is the wave speed, and $\xi = x - ct$. I obtained this equation by starting from Johnson's dimensionless equation (3.28) on page 208 and working backwards to untangle the scales that he had used to non-dimensionalize.
Alternatively, when the first term is expressed as a spatial derivative, the dimensional KdV equation is
\begin{equation}
2 \eta_x + \frac{3}{h_0} \eta \eta_{\xi} + \frac{h_0^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}
Notice the only difference is the absence of the wavespeed $c$ and the first term is now a spatial derivative.

@@

Let's establish some basic scales and notation:
% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table.}
\label{scale_table}
\begin{tabular}{l l l l}
\hline Description & Notation & Upstream  value \,\, & Downstream value \\
\hline
Characteristic wave amplitude	& $a = \etastd = \mean{\eta^2}^{1/2}$	& varied	& varied \\
Peak forcing frequency		& $f_p$				& 2 Hz		& N/A \\
Depth					& $h^{\pm}$			& 12.5 cm		& 3 cm \\
Peak wavelength			& $\lambda_p^{\pm}$	& 38 cm		& 25 cm \\
%Peak wavenumber			& $k_p$				& 0.17 rad/cm	& 0.25 rad/cm \\
%Dimensionless depth			& $k_p h$				& 2.1		& 0.76
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

\subsubsection{The variable-depth KdV equation}

We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
\begin{equation}
2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
\end{equation}
All variables above are dimensionless, and by working backwards I figured out how they relate to the original, dimensional variables. Below, the dimensional quantities are indicated with an overbar
\begin{align}
& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
\end{align}
Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the variable depth ratio.

There are a few important things to remember when we reformulate this equation
\begin{enumerate}
\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change. We have to do this so that the assumption of spatial periodicity will be reasonably satisfies.
\end{enumerate}
To achieve 2, we can be guided by the characteristic wavelength from shallow water theory
\begin{equation}
\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
\end{equation}
where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory). It is tempting to simply use the peak forcing frequency $f_p$ for $f_c$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$. Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments). Therefore, let us choose
\begin{equation}
\lambda_c = M \sqrt{gd} / f_p
\end{equation}
where $M$ is an integer (on the order of 10 or so).
